---
title: "Analysis of the Titanic Train dataset"
author: "Alice Otieno"
date: "7/11/2021"
output: html_document
---

### Import data
```{r}
train <- read.csv("C:/Users/tech/Desktop/train.csv")
```


### load the r packages that we'll need in our analysis
```{r}
library(tidyverse)
library(scales)

```

### Get a glance at the data,the top 10 records and bottom 6 records
```{r}
head(train,10)
tail(train) 
```

### See the variable names in the data set
```{r}
names(train)
```


### Structure of the data
```{r}
str(train)
```


### Summary statistics of the data set
```{r}
summary(train)
```

### Some variables in the data set are considered numeric and some as character but they are factors so we change the datatypes
```{r}
train$Survived <-as.factor(train$Survived)  
train$Pclass <-as.factor(train$Pclass)
train$Sex<-as.factor(train$Sex)
train$Embarked<-as.factor(train$Embarked)
str(train)
```



## Analysis and Visualization

This is the art of turning data into insight that can be easily interpreted as insight.
we'll check for survival patterns and check for factors that affected the same


### what was the survival rate?
We use a simple bar graph to demonstrate the survival rate
0 represents the passengers that did not survive, and 1 represents those that did survive. 

```{r}
ggplot(train,aes(x=Survived,fill=Survived))+geom_bar()
```
```{r}
prop.table(table(train$Survived))
```
From our bar graph we see that ~550 passengers did not survive, which is roughly 61% of the total, and ~340 (38%) survived


### Survival rate basis Gender
it is believed that in case of a rescue operation during disasters, women's safety is prioritized.Did the same happen?

```{r}
ggplot(train,aes(x=Sex,fill=Survived))+
  theme_bw()+
  geom_bar()+
  labs(y="Number of passengers",
  title="Survival Rate by Gender")

```

```{r}
prop.table(table(train$Survived,train$Sex))
```
The survival rate amongst women was significantly higher,compare to men.



### Survival rate basis class of tickets(Pclass)
There are 3levels of the ticket class,namely 1,2 and 3.

I presume the passengers in the pclass 1 were given first priority, is it the case?

```{r}
ggplot(train,aes(x=Pclass,fill=Survived))+
  theme_bw()+
  geom_bar()+
  labs(y="Number of passengers",
       title="Survival Rate basis Passenger Class")
```

We notice that over 50%of passengers used the 3rd class

About 77% of passengers in 1st class survived,about 44-50%from 2nd class and less than 25%in the 3rd class survived.




### Survival rate basis class of tickets(Pclass) and Gender

```{r}
ggplot(train,aes(x=Sex,fill=Survived))+
  theme_bw()+
  geom_bar()+
  labs(y="Number of Passengers",title="Titanic Survival Rate by Gender by Passenger class")+
  facet_wrap(~Pclass)
```
We can say that females were given first priority in the rescue plan. Number of female survivors was higher than male in all three passenger ticket classes.




### Survival rate basis age

We would want to know the impact of passenger's Age on survival rate

```{r}
ggplot(train,aes(x=Age, fill=Survived))+
    theme_bw()+
    geom_histogram(binwidth = 5)+
    labs(y="Number of passengers",x="Age", title = "survival rates basis Age")


```
looking at the age <10 years section in the graph,we see that the survival rate is low and drops above the age of 45.



### Survival rate basis Age ,Gender and Class of tickets

We also identify the survival pattern considering all the three variables

```{r}
ggplot(train,aes(x=Age,fill=Survived))+
  theme_bw()+
  facet_wrap(Pclass~Sex)+
  geom_histogram(bnwidth=5)+
  labs(y="survived",x="Age")
```
I
It is evident that the survival rate of children,across 1st and 2nd class was the highest.

Except for 1 girl child all children travelling 1st and 2nd class survived.

The survival rates were lowest for men traveling 3rd class.

**The End**
